# HMSC analysis

## Load data

```{r load_data_hmsc_origin_analysis}
load("data/data.Rdata")
if(!file.exists("hmsc_origin/fit_model1_250_10.Rdata")){
  download.file(url='https://sid.erda.dk/share_redirect/A6rm945NbY/hmsc_origin/fit_model1_250_10.Rdata',
              destfile='hmsc_origin/fit_model1_250_10.Rdata', method='curl')
}
load("hmsc_origin/fit_model1_250_10.Rdata")

# Select desired support threshold
support_threshold=0.9
negsupport_threshold=1-support_threshold
```

## Variance partitioning

```{r hmsc_variancepart, warning=FALSE, comments="", message=FALSE}

# Compute variance partitioning
varpart=computeVariancePartitioning(m)

varpart$vals %>%
   as.data.frame() %>%
   rownames_to_column(var="variable") %>%
   pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
   mutate(variable=factor(variable, levels=rev(c("origin","sex","logseqdepth","Random: location")))) %>%
   group_by(variable) %>%
   summarise(mean=mean(value)*100,sd=sd(value)*100) %>%
   tt()
```

```{r hmsc_varpart_plot, warning=FALSE, comments="", message=FALSE}
# Basal tree
varpart_tree <- genome_tree

#Varpart table
varpart_table <- varpart$vals %>%
   as.data.frame() %>%
   rownames_to_column(var="variable") %>%
   pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
   mutate(genome=factor(genome, levels=rev(varpart_tree$tip.label))) %>%
   mutate(variable=factor(variable, levels=rev(c("origin","sex","logseqdepth","Random: location"))))

#Phyla
phylum_groups <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__"))%>%
   right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% varpart_tree$tip.label) %>%
    arrange(match(genome, varpart_tree$tip.label)) %>%
    mutate(phylum = factor(phylum, levels = unique(phylum))) %>%
    column_to_rownames(var = "genome") %>%
    dplyr::select(phylum)

colors_alphabetic <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__"))%>%
  right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% varpart_tree$tip.label) %>%
    arrange(match(genome, varpart_tree$tip.label)) %>%
     dplyr::select(phylum, colors) %>%
    unique() %>%
    arrange(phylum) %>%
    dplyr::select(colors) %>%
    pull()

# Basal ggtree
varpart_tree <- varpart_tree %>%
        force.ultrametric(.,method="extend") %>%
        ggtree(., size = 0.3)

# Add phylum colors next to the tree tips
varpart_tree <- gheatmap(varpart_tree, phylum_groups, offset=-0.2, width=0.1, colnames=FALSE) +
   scale_fill_manual(values=colors_alphabetic)+
      labs(fill="Phylum")

#Reset fill scale to use a different colour profile in the heatmap
varpart_tree <- varpart_tree + new_scale_fill()

# Add variance stacked barplot
vertical_tree <-  varpart_tree +
       scale_fill_manual(values=c("#506a96","#cccccc","#be3e2b","#f6de6c"))+
        geom_fruit(
             data=varpart_table,
             geom=geom_bar,
             mapping = aes(x=value, y=genome, fill=variable, group=variable),
             pwidth = 2,
             offset = 0.05,
             width= 1,
             orientation="y",
             stat="identity")+
      labs(fill="Variable")

vertical_tree
```

## Model fit

```{r model_fit, warning=FALSE, comments="", message=FALSE}
MFCV <- evaluateModelFit(hM=m, predY=cv)

mean(MFCV$R2, na.rm = TRUE)

genome_fit <- tibble(genome=m$spNames, r2 = MFCV[[2]])

genome_counts_filt %>% 
mutate_if(is.numeric, ~ . / sum(.)) %>% 
left_join(genome_fit, by="genome") %>% 
filter(r2>0.5) %>% 
select(-c(genome,r2)) %>% 
colSums() %>%  
hist(main="Predictive capacity")
```

```{r predictive_mags, warning=FALSE, comments="", message=FALSE}
var_pred_table <- tibble(mag=m$spNames,
       pred=MFCV$R2,
       var_pred=MFCV$R2 * varpart$vals[1,],
       support=getPostEstimate(hM=m, parName="Beta")$support %>% .[2,],
       estimate=getPostEstimate(hM=m, parName="Beta")$mean %>% .[2,]) %>%
  mutate(enrichment=ifelse(support>=support_threshold,"Feral","Neutral")) %>% 
  mutate(enrichment=ifelse(support<=negsupport_threshold,"Domestic",enrichment))

var_pred_table %>% 
  ggplot(aes(x=estimate,y=var_pred, color=enrichment))+
    geom_point()+
    scale_color_manual(values=c("#bd70ae","#949293","#ebe8e8"))+
    geom_hline(yintercept=0.005, linetype="dashed")+
    theme_minimal()

predictive_mags <- var_pred_table %>% 
  filter(var_pred>=0.005) %>% 
  pull(mag)
```

## Posterior estimates

```{r hmsc_postestimates, warning=FALSE, comments="", message=FALSE}
# Basal tree
postestimates_tree <- genome_tree

# Posterior estimate table
post_beta <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
    mutate(genome=factor(genome, levels=rev(postestimates_tree$tip.label))) %>%
    mutate(value = case_when(
          value >= support_threshold ~ "Positive",
          value <= negsupport_threshold ~ "Negative",
          TRUE ~ "Neutral")) %>%
    mutate(value=factor(value, levels=c("Positive","Neutral","Negative"))) %>%
    pivot_wider(names_from = variable, values_from = value) %>%
    #select(genome,sp_vulgaris,area_semi,area_urban,sp_vulgarisxarea_semi,sp_vulgarisxarea_urban,season_spring,season_winter,sp_vulgarisxseason_spring,sp_vulgarisxseason_winter) %>%
    column_to_rownames(var="genome")

# Genome-specific attributes
genome_depth <- genome_counts_filt %>%
    mutate_if(is.numeric, ~ . / sum(.)) %>%
    rowwise() %>%
    mutate(depth = mean(c_across(where(is.numeric)), na.rm = TRUE)) %>%
    ungroup() %>%
    select(genome,depth)

#Phylums
phylum_groups <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__")) %>%
    right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% postestimates_tree$tip.label) %>%
    arrange(match(genome, postestimates_tree$tip.label)) %>%
    mutate(phylum = factor(phylum, levels = unique(phylum))) %>%
    column_to_rownames(var = "genome") %>%
    dplyr::select(phylum)

colors_alphabetic <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__")) %>%
    right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% postestimates_tree$tip.label) %>%
    arrange(match(genome, postestimates_tree$tip.label)) %>%
     dplyr::select(phylum, colors) %>%
    unique() %>%
    arrange(phylum) %>%
    dplyr::select(colors) %>%
    pull()

# Basal ggtree
postestimates_tree <- postestimates_tree %>%
        force.ultrametric(.,method="extend") %>%
        ggtree(., size = 0.3)

#Add phylum colors next to the tree tips
postestimates_tree <- gheatmap(postestimates_tree, phylum_groups, offset=-0.2, width=0.1, colnames=FALSE) +
      scale_fill_manual(values=colors_alphabetic)+
      labs(fill="Phylum")

#Reset fill scale to use a different colour profile in the heatmap
postestimates_tree <- postestimates_tree + new_scale_fill()

# Add posterior significant heatmap

postestimates_tree <- gheatmap(postestimates_tree, post_beta, offset=0, width=0.5, colnames=TRUE, colnames_position="top",colnames_angle=90, colnames_offset_y=1, hjust=0) +
        scale_fill_manual(values=c("#be3e2b","#f4f4f4","#b2b530"))+
        labs(fill="Trend")


postestimates_tree +
        vexpand(.25, 1) # expand top 
```
### Association with cat origin

```{r hmsc_support_plot_origin, warning=FALSE, comments="", message=FALSE}
estimate <- getPostEstimate(hM=m, parName="Beta")$mean %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    filter(variable=="originFeral") %>%
    pivot_longer(!variable, names_to = "genome", values_to = "mean") %>%
    dplyr::select(genome,mean)

support <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    filter(variable=="originFeral") %>%
    pivot_longer(!variable, names_to = "genome", values_to = "support") %>%
    dplyr::select(genome,support)

#pdf("figures/sig_Mags_violin.pdf",width=10, height=8)
inner_join(estimate,support,by=join_by(genome==genome)) %>%
    mutate(significance=ifelse(support >= 0.9 | support <= 0.1,1,0)) %>%
    mutate(support=ifelse(mean<0,1-support,support)) %>%
    left_join(genome_metadata, by = join_by(genome == genome)) %>%
    mutate(phylum = ifelse(support > 0.9, phylum, NA)) %>%
    ggplot(aes(x=mean,y=support,color=phylum))+
      geom_point(alpha=0.7, shape=16, size=3)+
      scale_color_manual(values = phylum_colors) +
      geom_vline(xintercept = 0) +
      xlim(c(-0.4,0.4)) +
      labs(x = "Beta regression coefficient", y = "Posterior probability") +
      theme_minimal()
```

```{r taxonomy_origin_all, warning=FALSE, comments="", message=FALSE}
estimate_quantiles <- getPostEstimate(hM=m, parName="Beta", q=c(0.1,0.9))$q

map_dfr(1:229, ~ {
  tibble(
    q10 = estimate_quantiles[1, 2, .x], # Extract 10% quantile
    q90 = estimate_quantiles[2, 2, .x]  # Extract 90% quantile
  )
}) %>% mutate(genome=estimate$genome) %>% 
  right_join(estimate, by="genome") %>% 
  left_join(genome_metadata, by="genome") %>% 
  filter(q10>0 | q90<0) %>% 
  ggplot(aes(x=mean, y=fct_reorder(genome, mean), xmin=q10, xmax=q90, color=phylum)) +
      geom_point() +
      geom_errorbar() +
      #xlim(c(-0.03,0.04)) +
      geom_vline(xintercept=0) +
      scale_color_manual(values = phylum_colors) +
      theme_minimal() +
      labs(x="Regression coefficient",y="Functional trait") +
      theme(legend.position="none")

```

```{r taxonomy_origin_top, warning=FALSE, comments="", message=FALSE}
estimate_top <- estimate %>% 
  arrange(desc(mean)) %>%                
  slice_head(n = 15) %>%                      
  bind_rows(                                  
    estimate %>%
      arrange(mean) %>%                      
      slice_head(n = 15))

estimate_top_significant <-map_dfr(1:229, ~ {
  tibble(
    q10 = estimate_quantiles[1, 2, .x], # Extract 10% quantile
    q90 = estimate_quantiles[2, 2, .x]  # Extract 90% quantile
  )
}) %>% mutate(genome=estimate$genome) %>% 
  right_join(estimate_top, by="genome") %>% 
  left_join(genome_metadata, by="genome") %>% 
  filter(q10>0 | q90<0) %>%
  arrange(mean)

estimate_top_significant %>% 
  ggplot(aes(x=mean, y=fct_reorder(genome, mean), xmin=q10, xmax=q90, color=phylum)) +
      geom_point() +
      geom_errorbar() +
      scale_y_discrete(labels = estimate_top_significant$genus) +
      geom_vline(xintercept=0) +
      scale_color_manual(values = phylum_colors[-c(5,7,9)]) +
      theme_minimal() +
      labs(x="Regression coefficient",y="Functional trait") +
      theme(legend.position="none")

```

#### Associated with domestic cats

```{r hmsc_positive_origin, warning=FALSE, comments="", message=FALSE}
origin_domestic <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
    mutate(trend = case_when(
          value >= support_threshold ~ "Positive",
          value <= negsupport_threshold ~ "Negative",
          TRUE ~ "Neutral")) %>%
    filter(variable=="originFeral", trend=="Negative") %>%
    arrange(-value) %>%
    left_join(genome_metadata,by=join_by(genome==genome)) %>%
    dplyr::select(genome,phylum,class,order,family,species,value) %>%
    arrange(phylum, class, family, species)

origin_domestic %>%
  paged_table()
```

#### Associated with feral cats

```{r hmsc_negative_origin, warning=FALSE, comments="", message=FALSE}
origin_feral <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
    mutate(trend = case_when(
          value >= support_threshold ~ "Positive",
          value <= negsupport_threshold ~ "Negative",
          TRUE ~ "Neutral")) %>%
    filter(variable=="originFeral", trend=="Positive") %>%
    arrange(value) %>%
    left_join(genome_metadata,by=join_by(genome==genome)) %>%
    dplyr::select(genome,phylum,class,order,family,species,value) %>%
    arrange(phylum,class,family,species)

origin_feral %>%
  paged_table()

```

## Phylogenetic signal

```{r phylogenetic_signal, warning=FALSE, comments="", message=FALSE}
mpost <- convertToCodaObject(m)
quantile(unlist(mpost$Rho), probs = c(.05,.5,.95))
```

## Microbiome composition predictions

Support values of the cat origin variable per genome are interpreted as enrichment for domestic or feral cats.

```{r hmsc_origin_enrichment, warning=FALSE, echo=FALSE, comments="", message=FALSE}
origin_enrichment <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
    mutate(enrichment = case_when(
          value >= support_threshold ~ "Feral",
          value <= negsupport_threshold ~ "Domestic",
          TRUE ~ "Neutral")) %>%
    filter(variable=="originFeral") %>%
    select(genome, enrichment)
```

The most likely microbiome compositions are predicted for domestic and feral cats based on the hmsc models

```{r prediction, warning=FALSE, echo=FALSE, comments="", message=FALSE}
gradient = c("domestic","feral")
gradientlength = length(gradient)

prediction <- constructGradient(m, 
                      focalVariable = "origin", 
                      non.focalVariables = 1) %>%
            predict(m, Gradient = ., expected = TRUE) %>%
            do.call(rbind,.) %>%
            as.data.frame() %>%
            mutate(origin=rep(gradient,1000)) %>%
            pivot_longer(!origin,names_to = "genome", values_to = "value")
```

Log-abundance differences between domestic and feral cats are plotted for genomes found enriched for either origin.

```{r sig_mag_plot, comment="", echo=FALSE, message=FALSE, warning=FALSE, fig.height=8, fig.width=10, fig.fullwidth=TRUE}
phylum_groups <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__")) %>%
    right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% genome_tree$tip.label) %>%
    arrange(match(genome, genome_tree$tip.label)) %>%
    mutate(phylum = factor(phylum, levels = unique(phylum))) %>%
    column_to_rownames(var = "genome") %>%
    dplyr::select(phylum)

colors_alphabetic <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__")) %>%
    right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% genome_tree$tip.label) %>%
    arrange(match(genome, genome_tree$tip.label)) %>%
     dplyr::select(phylum, colors) %>%
    unique() %>%
    arrange(phylum) %>%
    dplyr::select(colors) %>%
    pull()

genome_metadata_clean <- genome_metadata %>%  
    mutate(family = coalesce(family, paste('Unclassified', order)),
           genus = coalesce(genus, 
                              if_else(grepl('^Unclassified', family),
                                      family, paste('Unclassified', family))))

prediction %>%
  left_join(origin_enrichment,by="genome") %>% 
      filter(enrichment != "Neutral") %>% 
      pivot_wider(names_from = origin, values_from = value) %>%
      unnest(c(domestic, feral)) %>%
      mutate(diff_dom.feral = exp(feral)/exp(domestic)) %>%
      select(genome, diff_dom.feral) %>%
      left_join(genome_metadata_clean, by="genome") %>%
      mutate(genome= factor(genome, levels=genome_tree$tip.label)) %>%
      ggplot(., aes(y=forcats::fct_reorder(genus,diff_dom.feral), x=diff_dom.feral, fill=phylum, color=phylum)) +
            geom_boxplot(outlier.shape = NA) +
            scale_color_manual(values=phylum_colors)+
            scale_fill_manual(values=phylum_colors)+
          geom_vline(xintercept = 1)+
          coord_trans(x="log")+
          theme_minimal()+
  labs(x = "Domestic-to-feral abundance ratio", y = "Genera")

```

## Functional predictions

### Element level

```{r hmsc_origin_element, warning=FALSE, comments="", message=FALSE}
elements_table <- genome_gifts %>%
    to.elements(., GIFT_db) %>%
    as.data.frame() %>%
    .[predictive_mags,] #keep only predictive mags

#Calculate community-level functional traits for predicted communities
community_elements <- prediction %>%
  filter(genome %in% predictive_mags) %>% #keep only predictive mags
  group_by(origin, genome) %>%
  mutate(row_id = row_number()) %>%
  pivot_wider(names_from = genome, values_from = value) %>%
  ungroup() %>%
  group_split(row_id) %>%
  as.list() %>%
  lapply(., FUN = function(x){x %>%
    dplyr::select(-row_id) %>%
    column_to_rownames(var = "origin") %>%
    as.data.frame() %>%
    exp() %>%
    t() %>%
    tss() %>%
    to.community(elements_table,.,GIFT_db) %>% 
    as.data.frame() %>%
    rownames_to_column(var="origin")
   })

element_predictions <- map_dfr(community_elements, function(df) {
  domestic_values <- df %>% filter(origin == "domestic") %>% select(-origin)
  feral_values <- df %>% filter(origin == "feral") %>% select(-origin)
  domestic_values - feral_values
}) %>%
  mutate(iteration=c(1:1000)) %>% 
  pivot_longer(!iteration,names_to="trait",values_to="value") %>%
      group_by(trait) %>%
      summarise(mean=mean(value),
        p1 = quantile(value, probs = 0.1),
        p9 = quantile(value, probs = 0.9),
        positive_support = sum(value > 0)/1000,
        negative_support = sum(value < 0)/1000) %>%
      arrange(-positive_support)
```

#### Positively associated functions at element level

```{r hmsc_origin_element_significant, warning=FALSE, comments="", message=FALSE}
# Positively associated

unique_funct_db<- GIFT_db[c(2,4,5,6)] %>% 
  distinct(Code_element, .keep_all = TRUE)

element_predictions %>%
  filter(mean >0) %>%
  arrange(-positive_support) %>%
  filter(positive_support>=0.9) %>%
  left_join(unique_funct_db, by = join_by(trait == Code_element))%>%
  arrange(Domain,Function)%>%
  paged_table()

```

#### Negatively associated functions at element level

```{r hmsc_origin_element_significant_neg, warning=FALSE, comments="", message=FALSE}
element_predictions %>%
  filter(mean <0) %>%
  arrange(-negative_support) %>%
  filter(negative_support>=0.9) %>%
  left_join(unique_funct_db, by = join_by(trait == Code_element))%>%
  arrange(Domain,Function)%>%
  paged_table()
```

```{r hmsc_origin_significant_plot, warning=FALSE, comments="", message=FALSE, fig.height=7, fig.width=10, fig.fullwidth=TRUE}
positive_origin <- element_predictions %>%
  filter(mean >0) %>%
  arrange(mean) %>%
  filter(positive_support>=0.9) %>%
  dplyr::select(-negative_support) %>%
  rename(support=positive_support)

negative_origin <- element_predictions %>%
  filter(mean <0) %>%
  arrange(mean) %>%
  filter(negative_support>=0.9) %>%
  dplyr::select(-positive_support) %>%
  rename(support=negative_support)

all_elements <- bind_rows(positive_origin,negative_origin) %>%
  left_join(GIFT_db,by=join_by(trait==Code_element)) %>%
  mutate(trait=factor(trait,levels=c(rev(positive_origin$trait),rev(negative_origin$trait)))) %>%
  mutate(Code_function=factor(Code_function)) %>%
  mutate(element_legend=str_c(trait," - ",Element)) %>%
  mutate(function_legend=str_c(Code_function," - ",Function)) %>%
  select(trait,mean,p1,p9,element_legend,function_legend) %>% 
  unique()

gift_colors <- read_tsv("data/gift_colors.tsv") %>% 
  mutate(legend=str_c(Code_function," - ",Function))  %>% 
  filter(legend %in% all_elements$function_legend)

all_elements %>%
  ggplot(aes(x=-mean, y=fct_reorder(element_legend, -mean), xmin=-p1, xmax=-p9, color=function_legend)) +
      geom_point() +
      geom_errorbar() +
      xlim(c(-0.04,0.12)) +
      geom_vline(xintercept=0) +
      scale_color_manual(values = gift_colors$Color) +
      theme_minimal() +
      labs(x="Regression coefficient",y="Functional trait") +
    theme(legend.position="none")
```

### Function level

```{r hmsc_origin_function_predictions, warning=FALSE, comments="", message=FALSE}
functions_table <- elements_table %>%
    to.functions(., GIFT_db) %>%
    as.data.frame()

community_functions <- prediction %>%
  filter(genome %in% predictive_mags) %>% #keep only predictive mags
  group_by(origin, genome) %>%
  mutate(row_id = row_number()) %>%
  pivot_wider(names_from = genome, values_from = value) %>%
  ungroup() %>%
  group_split(row_id) %>%
  as.list() %>%
  lapply(., FUN = function(x){x %>%
    dplyr::select(-row_id) %>%
    column_to_rownames(var = "origin") %>%
    as.data.frame() %>%
    exp() %>%
    t() %>%
    tss() %>%
    to.community(functions_table,.,GIFT_db) %>% 
    as.data.frame() %>%
    rownames_to_column(var="origin")
   })
```

```{r hmsc_origin_function_predictions_model, warning=FALSE, comments="", message=FALSE}
function_predictions <- map_dfr(community_functions, function(df) {
  domestic_values <- df %>% filter(origin == "domestic") %>% select(-origin)
  feral_values <- df %>% filter(origin == "feral") %>% select(-origin)
  domestic_values - feral_values
}) %>%
  mutate(iteration=c(1:1000)) %>% 
  pivot_longer(!iteration,names_to="trait",values_to="value") %>%
      group_by(trait) %>%
      summarise(mean=mean(value),
        p1 = quantile(value, probs = 0.1),
        p9 = quantile(value, probs = 0.9),
        positive_support = sum(value > 0)/1000,
        negative_support = sum(value < 0)/1000) %>%
      arrange(-positive_support)

# Positively associated
function_predictions %>%
  filter(mean >0) %>%
  arrange(-positive_support) %>%
  paged_table()

# Negatively associated
function_predictions %>%
  filter(mean <0) %>%
  arrange(-negative_support) %>%
  paged_table()
```


```{r hmsc_origin_function_significant_plot, warning=FALSE, comments="", message=FALSE, fig.height=7, fig.width=10, fig.fullwidth=TRUE}

positive <- function_predictions %>%
  filter(mean >0) %>%
  arrange(mean) %>%
  filter(positive_support>=0.9) %>%
  dplyr::select(-negative_support) %>%
  rename(support=positive_support)

negative <- function_predictions %>%
  filter(mean <0) %>%
  arrange(mean) %>%
  filter(negative_support>=0.9) %>%
  dplyr::select(-positive_support) %>%
  rename(support=negative_support)

all_functions <- function_predictions %>%
  left_join(GIFT_db,by=join_by(trait==Code_function)) %>%
  mutate(trait=factor(trait)) %>%
  mutate(function_legend=str_c(trait," - ",Function)) %>%
  select(trait,mean,p1,p9,function_legend) %>% 
  unique()

gift_colors <- read_tsv("data/gift_colors.tsv") %>% 
  mutate(legend=str_c(Code_function," - ",Function))  %>% 
  filter(legend %in% all_functions$function_legend)

all_functions %>%
  ggplot(aes(x=-mean, y=fct_reorder(function_legend, -mean), xmin=-p1, xmax=-p9, color=function_legend)) +
      geom_point() +
      geom_errorbar() +
      xlim(c(-0.05,0.12)) +
      geom_vline(xintercept=0) +
      scale_color_manual(values = gift_colors$Color) +
      theme_minimal() +
      labs(x="Regression coefficient",y="Functional trait") +
      guides(col = guide_legend(ncol = 1))
```

