# HMSC analysis

## Load data

```{r load_data_hmsc_behaviour_analysis}
load("data/data.Rdata")
if(!file.exists("hmsc_behaviour/fit_model1_250_10.Rdata")){
  download.file(url='https://sid.erda.dk/share_redirect/A6rm945NbY/hmsc_behaviour/fit_model1_250_10.Rdata',
              destfile='hmsc/fit_model1_250_10.Rdata', method='curl')
}
load("hmsc_behaviour/fit_model1_250_10.Rdata")

# Select desired support threshold
support_threshold=0.9
negsupport_threshold=1-support_threshold
```

## Variance partitioning

```{r hmsc_behaviour_variancepart, warning=FALSE, comments="", message=FALSE}

# Compute variance partitioning
varpart=computeVariancePartitioning(m)

varpart$vals %>%
   as.data.frame() %>%
   rownames_to_column(var="variable") %>%
   pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
   mutate(variable=factor(variable, levels=rev(c("behaviour","sex","logseqdepth","Random: location")))) %>%
   group_by(variable) %>%
   summarise(mean=mean(value)*100,sd=sd(value)*100) %>%
   tt()
```

```{r hmsc_behaviour_varpart_plot, warning=FALSE, comments="", message=FALSE}
# Basal tree
varpart_tree <- genome_tree

#Varpart table
varpart_table <- varpart$vals %>%
   as.data.frame() %>%
   rownames_to_column(var="variable") %>%
   pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
   mutate(genome=factor(genome, levels=rev(varpart_tree$tip.label))) %>%
   mutate(variable=factor(variable, levels=rev(c("behaviour","sex","logseqdepth","Random: location"))))

#Phyla
phylum_groups <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__"))%>%
   right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% varpart_tree$tip.label) %>%
    arrange(match(genome, varpart_tree$tip.label)) %>%
    mutate(phylum = factor(phylum, levels = unique(phylum))) %>%
    column_to_rownames(var = "genome") %>%
    dplyr::select(phylum)

colors_alphabetic <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__"))%>%
  right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% varpart_tree$tip.label) %>%
    arrange(match(genome, varpart_tree$tip.label)) %>%
     dplyr::select(phylum, colors) %>%
    unique() %>%
    arrange(phylum) %>%
    dplyr::select(colors) %>%
    pull()

# Basal ggtree
varpart_tree <- varpart_tree %>%
        force.ultrametric(.,method="extend") %>%
        ggtree(., size = 0.3)

# Add phylum colors next to the tree tips
varpart_tree <- gheatmap(varpart_tree, phylum_groups, offset=-0.2, width=0.1, colnames=FALSE) +
   scale_fill_manual(values=colors_alphabetic)+
      labs(fill="Phylum")

#Reset fill scale to use a different colour profile in the heatmap
varpart_tree <- varpart_tree + new_scale_fill()

# Add variance stacked barplot
vertical_tree <-  varpart_tree +
       scale_fill_manual(values=c("#506a96","#cccccc","#be3e2b","#f6de6c"))+
        geom_fruit(
             data=varpart_table,
             geom=geom_bar,
             mapping = aes(x=value, y=genome, fill=variable, group=variable),
             pwidth = 2,
             offset = 0.05,
             width= 1,
             orientation="y",
             stat="identity")+
      labs(fill="Variable")

vertical_tree
```

## Model fit

```{r hmsc_bahaviour_model_fit, warning=FALSE, comments="", message=FALSE}
MFCV <- evaluateModelFit(hM=m, predY=cv)

mean(MFCV$R2, na.rm = TRUE)

genome_fit <- tibble(genome=m$spNames, r2 = MFCV[[2]])

genome_counts_filt %>% 
mutate_if(is.numeric, ~ . / sum(.)) %>% 
left_join(genome_fit, by="genome") %>% 
filter(r2>0.5) %>% 
select(-c(genome,r2)) %>% 
colSums() %>%  
hist(main="Predictive capacity")
```


```{r hmsc_behaviour_predictive_mags, warning=FALSE, comments="", message=FALSE}
var_pred_table <- tibble(mag=m$spNames,
       pred=MFCV$R2,
       var_pred=MFCV$R2 * varpart$vals[1,],
       support=getPostEstimate(hM=m, parName="Beta")$support %>% .[2,],
       estimate=getPostEstimate(hM=m, parName="Beta")$mean %>% .[2,]) %>%
  mutate(enrichment=ifelse(support>=support_threshold,"Feral","Neutral")) %>% 
  mutate(enrichment=ifelse(support<=negsupport_threshold,"Domestic",enrichment))

var_pred_table %>% 
  ggplot(aes(x=estimate,y=var_pred, color=enrichment))+
    geom_point()+
    scale_color_manual(values=c("#bd70ae","#949293","#ebe8e8"))+
    geom_hline(yintercept=0.005, linetype="dashed")+
    theme_minimal()

predictive_mags <- var_pred_table %>% 
  filter(var_pred>=0.005) %>% 
  pull(mag)
```

## Posterior estimates

```{r hmsc_behaviour_postestimates, warning=FALSE, comments="", message=FALSE}
# Basal tree
postestimates_tree <- genome_tree

# Posterior estimate table
post_beta <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
    mutate(genome=factor(genome, levels=rev(postestimates_tree$tip.label))) %>%
    mutate(value = case_when(
          value >= support_threshold ~ "Positive",
          value <= negsupport_threshold ~ "Negative",
          TRUE ~ "Neutral")) %>%
    mutate(value=factor(value, levels=c("Positive","Neutral","Negative"))) %>%
    pivot_wider(names_from = variable, values_from = value) %>%
    #select(genome,sp_vulgaris,area_semi,area_urban,sp_vulgarisxarea_semi,sp_vulgarisxarea_urban,season_spring,season_winter,sp_vulgarisxseason_spring,sp_vulgarisxseason_winter) %>%
    column_to_rownames(var="genome")

# Genome-specific attributes
genome_depth <- genome_counts_filt %>%
    mutate_if(is.numeric, ~ . / sum(.)) %>%
    rowwise() %>%
    mutate(depth = mean(c_across(where(is.numeric)), na.rm = TRUE)) %>%
    ungroup() %>%
    select(genome,depth)

#Phylums
phylum_groups <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__")) %>%
    right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% postestimates_tree$tip.label) %>%
    arrange(match(genome, postestimates_tree$tip.label)) %>%
    mutate(phylum = factor(phylum, levels = unique(phylum))) %>%
    column_to_rownames(var = "genome") %>%
    dplyr::select(phylum)

colors_alphabetic <- read_tsv("https://raw.githubusercontent.com/earthhologenome/EHI_taxonomy_colour/main/ehi_phylum_colors.tsv") %>%
    mutate(phylum=str_remove_all(phylum, "p__")) %>%
    right_join(genome_metadata, by=join_by(phylum == phylum)) %>%
    filter(genome %in% postestimates_tree$tip.label) %>%
    arrange(match(genome, postestimates_tree$tip.label)) %>%
     dplyr::select(phylum, colors) %>%
    unique() %>%
    arrange(phylum) %>%
    dplyr::select(colors) %>%
    pull()

# Basal ggtree
postestimates_tree <- postestimates_tree %>%
        force.ultrametric(.,method="extend") %>%
        ggtree(., size = 0.3)

#Add phylum colors next to the tree tips
postestimates_tree <- gheatmap(postestimates_tree, phylum_groups, offset=-0.2, width=0.1, colnames=FALSE) +
      scale_fill_manual(values=colors_alphabetic)+
      labs(fill="Phylum")

#Reset fill scale to use a different colour profile in the heatmap
postestimates_tree <- postestimates_tree + new_scale_fill()

# Add posterior significant heatmap

postestimates_tree <- gheatmap(postestimates_tree, post_beta, offset=0, width=0.5, colnames=TRUE, colnames_position="top",colnames_angle=90, colnames_offset_y=1, hjust=0) +
        scale_fill_manual(values=c("#be3e2b","#f4f4f4","#b2b530"))+
        labs(fill="Trend")


postestimates_tree +
        vexpand(.25, 1) # expand top 
```
### Association with cat behaviour

```{r hmsc_support_plot_behaviour, warning=FALSE, comments="", message=FALSE}
estimate <- getPostEstimate(hM=m, parName="Beta")$mean %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    filter(variable=="behaviour") %>%
    pivot_longer(!variable, names_to = "genome", values_to = "mean") %>%
    dplyr::select(genome,mean)

support <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    filter(variable=="behaviour") %>%
    pivot_longer(!variable, names_to = "genome", values_to = "support") %>%
    dplyr::select(genome,support)

#pdf("figures/sig_Mags_violin.pdf",width=10, height=8)
inner_join(estimate,support,by=join_by(genome==genome)) %>%
    mutate(significance=ifelse(support >= 0.9 | support <= 0.1,1,0)) %>%
    mutate(support=ifelse(mean<0,1-support,support)) %>%
    left_join(genome_metadata, by = join_by(genome == genome)) %>%
    mutate(phylum = ifelse(support > 0.9, phylum, NA)) %>%
    ggplot(aes(x=mean,y=support,color=phylum))+
      geom_point(alpha=0.7, shape=16, size=3)+
      scale_color_manual(values = phylum_colors) +
      geom_vline(xintercept = 0) +
      xlim(c(-0.4,0.4)) +
      labs(x = "Beta regression coefficient", y = "Posterior probability") +
      theme_minimal()
```

#### Associated with tame cats

```{r hmsc_positive_behaviour, warning=FALSE, comments="", message=FALSE}
behaviour_tame <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
    mutate(trend = case_when(
          value >= support_threshold ~ "Positive",
          value <= negsupport_threshold ~ "Negative",
          TRUE ~ "Neutral")) %>%
    filter(variable=="behaviour", trend=="Negative") %>%
    arrange(-value) %>%
    left_join(genome_metadata,by=join_by(genome==genome)) %>%
    dplyr::select(genome,phylum,class,order,family,species,value) %>%
    arrange(phylum, class, family, species)

behaviour_tame %>%
  arrange(value) %>% 
  paged_table()

```

#### Associated with aggresive cats

```{r hmsc_negative_behaviour, warning=FALSE, comments="", message=FALSE}
behaviour_aggresive <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
    mutate(trend = case_when(
          value >= support_threshold ~ "Positive",
          value <= negsupport_threshold ~ "Negative",
          TRUE ~ "Neutral")) %>%
    filter(variable=="behaviour", trend=="Positive") %>%
    arrange(value) %>%
    left_join(genome_metadata,by=join_by(genome==genome)) %>%
    dplyr::select(genome,phylum,class,order,family,species,value) %>%
    arrange(phylum,class,family,species)

behaviour_aggresive %>%
  arrange(-value) %>% 
  paged_table()

```
## Behaviour-origin congruence

```{r behaviour_origin_congruence, warning=FALSE, comments="", message=FALSE}

tame_domestic <- intersect(behaviour_tame$genome, origin_domestic$genome) %>% length()
tame_feral <- intersect(behaviour_tame$genome, origin_feral$genome) %>% length()
aggressive_domestic <- intersect(behaviour_aggresive$genome, origin_domestic$genome) %>% length()
aggressive_feral <- intersect(behaviour_aggresive$genome, origin_feral$genome) %>% length()

tibble(origin=c("domestic","feral"),aggressive=c(aggressive_domestic,aggressive_feral),tame=c(tame_domestic,tame_feral)) %>% 
  tt()

behaviour_tame %>% 
  filter(genome %in% intersect(behaviour_tame$genome, origin_domestic$genome)) %>%
  paged_table()

behaviour_aggresive %>% 
  filter(genome %in% intersect(behaviour_aggresive$genome, origin_domestic$genome)) %>%
  paged_table()

behaviour_aggresive %>% 
  filter(genome %in% intersect(behaviour_aggresive$genome, origin_feral$genome)) %>%
  paged_table()
```

## Domestic & aggressive cats

```{r domestic_aggressive cats, warning=FALSE, comments="", message=FALSE}
domestic_aggressive <- sample_metadata %>% 
        rowwise() %>%
        mutate(behaviour = sum(c_across(c(bites, hisses,retreats,avoidance)), na.rm = TRUE)/4) %>%
        ungroup() %>%
        filter(origin == "Domestic" & behaviour > 0) %>%
        pull(sample)

domestic_tame <- sample_metadata %>% 
        rowwise() %>%
        mutate(behaviour = sum(c_across(c(bites, hisses,retreats,avoidance)), na.rm = TRUE)/4) %>%
        ungroup() %>%
        filter(origin == "Domestic" & behaviour == 0) %>%
        pull(sample)

focal_genomes <- behaviour_aggresive %>% 
  filter(genome %in% intersect(behaviour_aggresive$genome, origin_domestic$genome))  %>% 
  filter(family == "Megasphaeraceae") %>% 
  pull(genome)


phylo_samples <- sample_metadata %>% 
    filter(sample %in% c(domestic_aggressive,domestic_tame)) %>% 
    mutate(behaviour = case_when(
    sample %in% domestic_aggressive ~ "aggresive",
    sample %in% domestic_tame ~ "tame",
    TRUE ~ NA_character_
  )) %>% 
    column_to_rownames("sample") %>% 
    sample_data()

phylo_genome <- genome_counts_filt %>% 
     filter(genome %in% focal_genomes) %>% 
     column_to_rownames("genome") %>% 
     mutate_all(~ replace(., . == 0, 0.00001)) %>%
      otu_table(., taxa_are_rows = TRUE)

phylo_taxonomy <- genome_metadata %>% 
     filter(genome %in% focal_genomes) %>% 
     mutate(genome2=genome) %>% #create a pseudo genome name column
     column_to_rownames("genome2") %>% 
     dplyr::select(division,phylum,class,order,family,genus,species,genome) %>%
     as.matrix() %>% 
     tax_table()

physeq_genome_filtered <- phyloseq(phylo_genome, phylo_taxonomy, phylo_samples)

ancom_rand_output = ancombc2(data = physeq_genome_filtered, 
                  assay_name = "counts",
                  tax_level = NULL, #change to agglomerate analysis to a higher taxonomic range
                  fix_formula = "behaviour", #fixed variable(s)
#                  rand_formula = "(1|Individual)",
                  p_adj_method = "holm", 
                  pseudo_sens = TRUE,
                  prv_cut = 0, 
                  s0_perc = 0,
                  group = NULL, 
                  struc_zero = FALSE, 
                  neg_lb = FALSE,
                  alpha = 0.05, 
                  n_cl = 2, 
                  verbose = TRUE,
                  global = FALSE, 
                  pairwise = FALSE, 
                  dunnet = FALSE, 
                  trend = FALSE,
                  iter_control = list(tol = 1e-5, max_iter = 20, verbose = FALSE),
                  em_control = list(tol = 1e-5, max_iter = 100),
                  lme_control = NULL,
                  mdfdr_control = list(fwer_ctrl_method = "holm", B = 100), 
                  trend_control = NULL)



genome_counts_filt %>%
  mutate_at(vars(-genome),~./sum(.)) %>% #apply TSS nornalisation
  pivot_longer(-genome, names_to = "sample", values_to = "count") %>% #reduce to minimum number of columns
  left_join(., genome_metadata, by = join_by(genome == genome)) %>% #append genome metadata
  left_join(., sample_metadata, by = join_by(sample == sample)) %>% 
  filter(genome %in% focal_genomes) %>% 
  filter(sample %in% c(domestic_aggressive,domestic_tame)) %>% 
  mutate(behaviour = case_when(
    sample %in% domestic_aggressive ~ "aggresive",
    sample %in% domestic_tame ~ "tame",
    TRUE ~ NA_character_
  )) %>% 
  ggplot(aes(x=behaviour,y=count, group=behaviour, color=genome))+
    geom_jitter()

```

## Phylogenetic signal

```{r hmsc_behaviour_phylogenetic_signal, warning=FALSE, comments="", message=FALSE}
mpost <- convertToCodaObject(m)
quantile(unlist(mpost$Rho), probs = c(.05,.5,.95))
```

## Microbiome composition predictions

Support values of the cat behaviour variable per genome are interpreted as enrichment for aggresuve or tame cats.

```{r hmsc_behaviour_enrichment, warning=FALSE, echo=FALSE, comments="", message=FALSE}
origin_enrichment <- getPostEstimate(hM=m, parName="Beta")$support %>%
    as.data.frame() %>%
    mutate(variable=m$covNames) %>%
    pivot_longer(!variable, names_to = "genome", values_to = "value") %>%
    mutate(enrichment = case_when(
          value >= support_threshold ~ "Aggresive",
          value <= negsupport_threshold ~ "Tame",
          TRUE ~ "Neutral")) %>%
    filter(variable=="behaviour") %>%
    select(genome, enrichment)
```

The most likely microbiome compositions are predicted for the gradient from tameness to aggressiveness based on the hmsc models

```{r hmsc_behaviour_prediction, warning=FALSE, echo=FALSE, comments="", message=FALSE}
gradient = seq(0,1,by=0.1)
gradientlength = length(gradient)

prediction <- constructGradient(m, 
                      focalVariable = "behaviour", 
                      non.focalVariables = 1, 
                      ngrid=gradientlength) %>%
            predict(m, Gradient = ., expected = TRUE) %>%
            do.call(rbind,.) %>%
            as.data.frame() %>%
            mutate(behaviour=rep(gradient,1000)) %>%
            pivot_longer(!behaviour,names_to = "genome", values_to = "value")
```

## Functional predictions

### Element level

```{r hmsc_behaviour_element, warning=FALSE, comments="", message=FALSE}
elements_table <- genome_gifts %>%
    to.elements(., GIFT_db) %>%
    as.data.frame() %>%
    .[predictive_mags,] #keep only predictive mags

#Calculate community-level functional traits for predicted communities
community_elements <- prediction %>%
  filter(genome %in% predictive_mags) %>% #keep only predictive mags
  group_by(behaviour, genome) %>%
  mutate(row_id = row_number()) %>%
  pivot_wider(names_from = genome, values_from = value) %>%
  ungroup() %>%
  group_split(row_id) %>%
  as.list() %>%
  lapply(., FUN = function(x){x %>%
    dplyr::select(-row_id) %>%
    column_to_rownames(var = "behaviour") %>%
    as.data.frame() %>%
    exp() %>%
    t() %>%
    tss() %>%
    to.community(elements_table,.,GIFT_db) %>% 
    as.data.frame() %>%
    rownames_to_column(var="behaviour")
   })

calculate_slope <- function(x) {
  lm_fit <- lm(unlist(x) ~ seq_along(unlist(x)))
  coef(lm_fit)[2]
}

element_predictions <- map_dfc(community_elements, function(df) {
  df %>%
        column_to_rownames(var = "behaviour") %>%
        t() %>%
        as.data.frame() %>%
        rowwise() %>%
        mutate(slope = calculate_slope(c_across(everything()))) %>%
        select(slope) }) %>%
      t() %>%
      as.data.frame() %>%
      set_names(colnames(community_elements[[1]])[-1]) %>%
      rownames_to_column(var="iteration") %>%
      pivot_longer(!iteration, names_to="trait",values_to="value") %>%
      group_by(trait) %>%
      summarise(mean=mean(value),
        p1 = quantile(value, probs = 0.1),
        p9 = quantile(value, probs = 0.9),
        positive_support = sum(value > 0)/1000,
        negative_support = sum(value < 0)/1000) %>%
      arrange(-positive_support)
```

#### Positive associated functions at element level

```{r hmsc_behaviour_element_significant, warning=FALSE, comments="", message=FALSE}
# Positively associated

unique_funct_db<- GIFT_db[c(2,4,5,6)] %>% 
  distinct(Code_element, .keep_all = TRUE)

element_predictions %>%
  filter(mean >0) %>%
  arrange(-positive_support) %>%
  filter(positive_support>=0.9) %>%
  left_join(unique_funct_db, by = join_by(trait == Code_element))%>%
  arrange(Domain,Function)%>%
  paged_table()

```

#### Negatively associated functions at element level

```{r hmsc_behaviour_element_significant_neg, warning=FALSE, comments="", message=FALSE}
element_predictions %>%
  filter(mean <0) %>%
  arrange(-negative_support) %>%
  filter(negative_support>=0.9) %>%
  left_join(unique_funct_db, by = join_by(trait == Code_element))%>%
  arrange(Domain,Function)%>%
  paged_table()
```

```{r hmsc_behaviour_significant_plot, warning=FALSE, comments="", message=FALSE, fig.height=7, fig.width=10, fig.fullwidth=TRUE}
positive <- element_predictions %>%
  filter(mean >0) %>%
  arrange(mean) %>%
  filter(positive_support>=0.9) %>%
  dplyr::select(-negative_support) %>%
  rename(support=positive_support)

negative <- element_predictions %>%
  filter(mean <0) %>%
  arrange(mean) %>%
  filter(negative_support>=0.9) %>%
  dplyr::select(-positive_support) %>%
  rename(support=negative_support)

all_elements <- bind_rows(positive,negative) %>%
  left_join(GIFT_db,by=join_by(trait==Code_element)) %>%
  mutate(trait=factor(trait,levels=c(rev(positive$trait),rev(negative$trait)))) %>%
  mutate(Code_function=factor(Code_function)) %>%
  mutate(element_legend=str_c(trait," - ",Element)) %>%
  mutate(function_legend=str_c(Code_function," - ",Function)) %>%
  select(trait,mean,p1,p9,element_legend,function_legend) %>% 
  unique()

gift_colors <- read_tsv("data/gift_colors.tsv") %>% 
  mutate(legend=str_c(Code_function," - ",Function))  %>% 
  filter(legend %in% all_elements$function_legend)

all_elements %>%
  ggplot(aes(x=mean, y=fct_reorder(element_legend, mean), xmin=p1, xmax=p9, color=function_legend)) +
      geom_point() +
      geom_errorbar() +
      xlim(c(-0.1,0.04)) +
      geom_vline(xintercept=0) +
      scale_color_manual(values = gift_colors$Color) +
      theme_minimal() +
      labs(x="Regression coefficient",y="Functional trait")
```

### Function level

```{r hmsc_behaviour_function_predictions, warning=FALSE, comments="", message=FALSE}
functions_table <- elements_table %>%
    to.functions(., GIFT_db) %>%
    as.data.frame()

community_functions <- prediction %>%
  filter(genome %in% predictive_mags) %>% #keep only predictive mags
  group_by(behaviour, genome) %>%
  mutate(row_id = row_number()) %>%
  pivot_wider(names_from = genome, values_from = value) %>%
  ungroup() %>%
  group_split(row_id) %>%
  as.list() %>%
  lapply(., FUN = function(x){x %>%
    dplyr::select(-row_id) %>%
    column_to_rownames(var = "behaviour") %>%
    as.data.frame() %>%
    exp() %>%
    t() %>%
    tss() %>%
    to.community(functions_table,.,GIFT_db) %>% 
    as.data.frame() %>%
    rownames_to_column(var="behaviour")
   })

calculate_slope <- function(x) {
  lm_fit <- lm(unlist(x) ~ seq_along(unlist(x)))
  coef(lm_fit)[2]
}

function_predictions <- map_dfc(community_functions, function(df) {
      df %>%
        column_to_rownames(var = "behaviour") %>%
        t() %>%
        as.data.frame() %>%
        rowwise() %>%
        mutate(slope = calculate_slope(c_across(everything()))) %>%
        select(slope) }) %>%
      t() %>%
      as.data.frame() %>%
      set_names(colnames(community_functions[[1]])[-1]) %>%
      rownames_to_column(var="iteration") %>%
      pivot_longer(!iteration, names_to="trait",values_to="value") %>%
      group_by(trait) %>%
      summarise(mean=mean(value),
        p1 = quantile(value, probs = 0.1),
        p9 = quantile(value, probs = 0.9),
        positive_support = sum(value > 0)/1000,
        negative_support = sum(value < 0)/1000) %>%
      arrange(-positive_support)
```

```{r hmsc_behaviour_function_significant_plot, warning=FALSE, comments="", message=FALSE, fig.height=7, fig.width=10, fig.fullwidth=TRUE}

positive <- function_predictions %>%
  filter(mean >0) %>%
  arrange(mean) %>%
  filter(positive_support>=0.9) %>%
  dplyr::select(-negative_support) %>%
  rename(support=positive_support)

negative <- function_predictions %>%
  filter(mean <0) %>%
  arrange(mean) %>%
  filter(negative_support>=0.9) %>%
  dplyr::select(-positive_support) %>%
  rename(support=negative_support)

all_functions <- function_predictions %>%
  left_join(GIFT_db,by=join_by(trait==Code_function)) %>%
  mutate(trait=factor(trait)) %>%
  mutate(function_legend=str_c(trait," - ",Function)) %>%
  select(trait,mean,p1,p9,function_legend) %>% 
  unique()

gift_colors <- read_tsv("data/gift_colors.tsv") %>% 
  mutate(legend=str_c(Code_function," - ",Function))  %>% 
  filter(legend %in% all_functions$function_legend)

all_functions %>%
  ggplot(aes(x=mean, y=fct_reorder(function_legend, mean), xmin=p1, xmax=p9, color=function_legend)) +
      geom_point() +
      geom_errorbar() +
      geom_vline(xintercept=0) +
      scale_color_manual(values = gift_colors$Color) +
      theme_minimal() +
      labs(x="Regression coefficient",y="Functional trait") +
      guides(col = guide_legend(ncol = 1))
```
